EBASE manuscript

### 2023-01-06 manu-draft.qmd
simulated apa fig and cross refs added

A preliminary assessment of the ability of EBASE to produce reasonable parameter estimates for $a$, $r$, and $b$ (@eq-ebase) was conducted using a simulated time series created with FWOXY. These assessments evaluated if 1) EBASE estimates similar values for the known parameters, and 2) how the estimated values change with different model configurations. The simulated time series used inputs from an approximate one year of continuous data from Apalachicola Bay, Florida.  The input data were 2012 observations from February 22nd to December 19th at the Cat Point water quality monitoring station (29.7021$^{\circ}$ N, -84.8802$^{\circ}$ W) and included 15 minute observations of dissolved oxygen (mg/L), water temperature (C), and salinity (psu).  Continuous meteorological observations from the nearby East Bay station (29.7909$^{\circ}$ N, -84.8834$^{\circ}$ W, ~10km north of Cat Point) were combined with the water quality data at the same time step and included air temperature (C), wind speed (m/s), and PAR (mmol/m$^2$ as total per 15 minute observation). An oxygen time series was generated from Fwoxy using the inputs from Apalachicola Bay, such that the simulated dissolved oxygen was created as close as possible to the observed given the inputs.  Known values for $a$ and $r$ parameters were returned for each time step of the simulated data, as well as all metabolic estimates for production, respiration, and gas exchange.  The only parameter that could not be simulated by Fwoxy was $b$, which was set to 0.251.

The simulated time series is shown in figure......

@@

The metabolic estimates returned from each analysis were compared to the known values to identify priors the model was most sensitive to and if the conclusions varied by optimization period.  Estimates for the modeled dissolved oxygen, gross production, respiration ($r$), gas exchange ($D$), and $a$ were compared to those from the simulated time series using coefficient of variation ($R^2$) and root mean square error ($RMSE$) summaries of the 1:1 comparison of the estimated vs simulated values.  For the one day optimization period, all parameter estimates returned at the time step of the input data (i.e., modeled dissolved oxygen, $D$, and $a$) were daily-averaged for comparison.  For the seven day optimization period, all parameter estimates and those from the known time series were weekly-averaged.  Estimates for the $b$ parameter were evaluated only by visual comparison because a single value was used in the simulated time series, preventing the estimate of quantitative comparisons.

@@

>>>>

A preliminary assessment of the ability of EBASE to produce reasonable parameter estimates for $a$, $r$, and $b$ (@eq-ebase) was conducted using a simulated time series created with FWOXY. These assessments evaluated if 1) EBASE estimates similar values for the known parameters, and 2) how the estimated values change with different model configurations. The simulated time series used inputs from an approximate one year of continuous data from Apalachicola Bay, Florida.  The input data were 2012 observations from February 22nd to December 19th at the Cat Point water quality monitoring station (29.7021$^{\circ}$ N, -84.8802$^{\circ}$ W) and included 15 minute observations of dissolved oxygen (mg/L), water temperature (C), and salinity (psu).  Continuous meteorological observations from the nearby East Bay station (29.7909$^{\circ}$ N, -84.8834$^{\circ}$ W, ~10km north of Cat Point) were combined with the water quality data at the same time step and included air temperature (C), wind speed (m/s), and PAR (mmol/m$^2$ as total per 15 minute observation). An oxygen time series was generated from Fwoxy using the inputs from Apalachicola Bay ([@fig-simapa]a), such that the simulated dissolved oxygen was created as close as possible to the observed given the inputs.  Known values for $a$ and $r$ parameters were returned for each time step of the simulated data, as well as all metabolic estimates for production, respiration, and gas exchange ([@fig-simapa]b and c).  The only parameter that could not be simulated by Fwoxy was $b$, which was set to 0.251.

@@

@@

![Simulated time series using the FWOXY model for one year of continuous data at Apalachicola Bay.  Simulated input to EBASE was the (a) dissolved oxygen time series, with results compared to the (b) simulated $a$ parameter and (c) simulatd metabolic estimates for production (P), respiration (R), and gas exchange (D).](figs/simapa.png){#fig-simapa}
